---
title: "Analyzing Suicide Rates Among Adolescents"
author: "Grayson Ellard, Keaton Staggs, Yingxuan Wang, Netra Krishnan"
date: "11/14/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Summary Paragraph
```{r, echo=FALSE, message=FALSE, warning=FALSE}
source("~/info201/project/project-Graydwarf77/source/summary_info.R")
```
Our project primarily uses data from the "crude suicide rates" data file. This data set contains `r summary_info$num_country_suicide_rate ` countries, missing about 12 from the total of 195 countries in the world. In this set, the mean suicide rate per 100,000 for adolescents between 10 and 19 is `r summary_info$mean_rate_country`. This means that around the entire world, about 4 adolescents die of suicide per 100,000. The median suicide rate per 100,000 in this data set is `r summary_info$median_rate_country`, which is slightly less than the mean. Overall, the highest suicide rate in this data set is `r summary_info$highest_rate_country`, which is over four times the mean. Our project also used the "Human resources" data file, which has information on mental health resources within many countries around the world. This data set contains `r summary_info$num_countries_human_resources` countries, which is considerably less than the suicide rates data set. 


## Table
```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.show='hide'}
source("~/info201/project/project-Graydwarf77/source/summary_table.R")
summary_table
```

This table displays the overall data found in our research and analysis of suicide rates in adolescents around the world. It includes data of the comparison between suicide rates in adolescents across countries, between male and female adolescents, and in comparison to the rate of mental health resources available across the world. This table is included to provide the ability to compare suicide rates in different contexts, through a holistic view of the different variables affecting suicide rates. Through the data in the table, it was found that each country found a higher suicide rate for male adolescents in comparison to female adolescents, except for a handful of outlier countries such as Bangladesh, Fiji, Guatemala, Indonesia, Lesotho, Maldives, Morocco, Myanmar, Nepal, Paraguay, Sri Lanka, Suriname, Timor-Leste, and Uzbekistan. By comparing this data with other summary data in the table, it can be found that a couple of these outlier countries have a lower human resources rate in comparison to their suicide rates (Nepal, Myanmar, and Bangladesh). However, most of the countries in this section have unavailable information relating to their human resources, so a conclusion can be hard to draw relating human resources to the difference in suicide rates between male and female adolescents. Besides this, the table also allows one to directly look at suicide rates side-by-side with human resources rates in each country, being able to visually look through data around the world and draw connections. By looking at the table, one can find countries like Norway, Panama, and Peru having extremely low suicide rates and extremely high human resources rates, and looking at the specific numbers and statistics to back it. The charts don't allow for this, since they are meant to be more visually pleasing, rather than allowing the viewer to hone in on the statistics.

## Chart One
```{r, echo=FALSE, message=FALSE, warning=FALSE}
source("~/info201/project/project-Graydwarf77/source/Chart1.R")
Chart_1
```

The purpose of this chart is to find the comparison between male and female suicide rates in adolescents around the world. If a major discrepancy is found, it can show how mental health resources could be allocated better towards certain groups moving forward. Based on this chart, it was found that through an average of suicide rates around the world, male adolescents had nearly twice the number of suicide rates than female adolescents. Mental health researchers can use this information to better inform their research towards treatment and better understanding the effects of suicidal ideation.


## Chart Two
```{r, echo=FALSE, message=FALSE, warning=FALSE}
source("~/info201/project/project-Graydwarf77/source/chart2.R")
chart_2
```

Geared toward our second research question, this data visualization displays the suicide rate for adolescents in the United States in comparison to various other countries throughout the world. The pie chart visualization was created to express the differing amount of adolescent suicides (per 100,000) for various nations, with a focus on how the United States compares to other countries. Revealed from the chart, the United States experiences 6.3 suicide deaths per 100,000 people. The highest amount of suicides is in India, with 11.2 suicides per 100,000 people and the lowest amount is 1.9 in Malaysia. From the data, we can observe that the United States has a high suicide rate among adolescents, being the second highest value within the selected countries.



## Chart Three
```{r, echo=FALSE, message=FALSE, warning=FALSE}
source("~/info201/project/project-Graydwarf77/source/Chart_3.R")
chart_three
```

This chart is aimed towards understanding how the number of mental health care workers in a country correlates to that country's adolescent suicide rate. We did this because it is important to understand how different resources and approaches may affect the suicide rate of adolescents in different countries. Interestingly, out of visualized countries (which were selected based on completeness of data and geological diversity), ones which have few mental health resources seem to have suicide rates that are on par or below that of countries that have high mental health resources. India is an outlier for the group, with a suicide rate of 11.2 per 100,000. That is nearly double the next closest, which is the United States at a suicide rate of 6.2 per 100,000.
