---
title: "Analyzing Suicide Rates Among Adolescents"
author: "Grayson Ellard, Keaton Staggs, Yingxuan Wang, Netra Krishnan"
date: "1/22/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Summary Paragraph

## Table

## Chart One
```{r, echo=FALSE, message=FALSE, warning=FALSE}

```
The purpose of this chart is to find the comparison between male and female suicide rates in adolescents around the world. If a major discrepancy is found, it can show how mental health resources could be allocated better towards certain groups moving forward. Based on this chart, it was found that through an average of suicide rates around the world, male adolescents had nearly twice the number of suicide rates than female adolescents. Mental health researchers can use this information to better inform their research towards treatment and better understanding the effects of suicidal ideation.


## Chart Two
```{r, echo=FALSE, message=FALSE, warning=FALSE}

```
Geared toward our second research question, this data visualization displays the suicide rate for adolescents in the United States in comparison to various other countries throughout the world. The pie chart visualization was created to express the differing amount of adolescent suicides (per 100,000) for various nations, with a focus on how the United States compares to other countries. Revealed from the chart, the United States experiences 6.3 suicide deaths per 100,000 people. The highest amount of suicides is in India, with 11.2 suicides per 100,000 people and the lowest amount is 1.9 in Malaysia. From the data, we can observe that the United States has a high suicide rate among adolescents, being the second highest value within the selected countries.



## Chart Three
```{r, echo=FALSE, message=FALSE, warning=FALSE}
source("~/info201/project/project-Graydwarf77/source/Chart_3.R")
chart_three
```

This chart is aimed towards understanding how the number of mental health care workers in a country correlates to that country's adolescent suicide rate. We did this because it is important to understand how different resources and approaches may affect the suicide rate of adolescents in different countries. Interestingly, out of visualized countries (which were selected based on completeness of data and geological diversity), ones which have few mental health resources seem to have suicide rates that are on par or below that of countries that have high mental health resources. India is an outlier for the group, with a suicide rate of 11.2 per 100,000. That is nearly double the next closest, which is the United States at a suicide rate of 6.2 per 100,000.